### 2025-07-05 202507_uai_puncs/arxiv/content.tex
y54wytr

Using the concept of partition circuits we construct positive operator circuits. Positive operator circuits can be thought of as generalizing circuit evaluations with probabilities to circuit evaluations with PSD matrices.
% Note, in the definition below we use $\Ocircuit_k$ instead of $\circuit_k$ to make this generalization explicit.
\begin{definition}[Positive Operator Circuit (Partition Circuit)]
	\label{def:poc}
	Let   $\xvars{=}\{\xvar_0,\dots ,\xvar_{\numvar{-}1}  \}$ be a set of discrete variables.
	We define an operator circuit as a partition circuit whose computation units take the following functional form:
	\begin{align}
		 & {\Ocircuit}_k(\xvars_k){=}
		\begin{cases}
			E_{\xvar_k}
			% A_{k} \times  e_{\xvar_k} \otimes  e^*_{\xvar_k} \times A_{k}^*,
			 & \text{if $k$ is leaf}
			\\
			\qop_k \Bigl(\Ocircuit_{k_l}(\xvars_{k_l}) \otimes \Ocircuit_{k_r}(\xvars_{k_r}) \Bigr)
			 & \text{else},
		\end{cases}
		\label{eq:poc:def}
	\end{align}
	where the  $E_{\xvar_k}$'s are quantum state matrices, and where the $\qop_k$ quantum operations.
\end{definition}
Note that using the Kronecker product between $\Ocircuit_{k_l}(\xvars_{k_l})$ and  $\Ocircuit_{k_r}(\xvars_{k_r})$ is a sensible choice as it describes the joint state of both subsystems.

@@

To show that PSD circuits are a special case of \puncs we now impose the following restriction on the quantum operations $\qop_k$:
\begin{align}
	\qop_k (\Ocircuit_{k_l} \otimes \Ocircuit_{k_l} )
	 & =
	\kraus_{k} \left( \Ocircuit_{k_l}  \otimes  \Ocircuit_{k_r} \right) \kraus_{k}^*
	\label{eq:pureinternal}
\end{align}
That is, we limit the quantum operation to having  only a single pair of Kraus operators. For the quantum operation to be unital we need to have $\kraus_{k} \kraus_{k}^*=\mathbb{1}$. That is, $\kraus_{k}$ has to be semi-unitary,

@@

Effectively, this would expand the research initiated by \citet{loconte2024subtractive} for mapping out the relationships between the different structured decomposable circuits. Ideally, one would like to create an analogue to \citeauthor{darwiche2002knowledge}' knowledge compilation map \citep{darwiche2002knowledge} but for positive operator circuits and relate it to the framework of algebraic model counting \citep{kimmig2017algebraic}.
This would also allow for establishing possible separations between sum-of-squares PCs~\citep{loconte2024subtractive}, product of monotonic by SOCS~\citep[Definition 5]{loconte2024subtractive}, and inception networks~\citep{wangrelationship}.
A first discussion on these issues can also be fund in \citep[Appendix A]{wangrelationship}.

>>>>

Using the concept of partition circuits we construct positive operator circuits. Positive operator circuits can be thought of as generalizing circuit evaluations with probabilities to circuit evaluations with PSD matrices.
% Note, in the definition below we use $\Ocircuit_k$ instead of $\circuit_k$ to make this generalization explicit.
\begin{definition}[Positive Operator Circuit (Partition Circuit)]
	\label{def:poc}
	Let   $\xvars{=}\{\xvar_0,\dots ,\xvar_{\numvar{-}1}  \}$ be a set of discrete variables.
	We define an operator circuit as a partition circuit whose computation units take the following functional form:
	\begin{align}
		 & {\Ocircuit}_k(\xvars_k){=}
		\begin{cases}
			E_{\xvar_k}
			% A_{k} \times  e_{\xvar_k} \otimes  e^*_{\xvar_k} \times A_{k}^*,
			 & \text{if $k$ is leaf}
			\\
			\qop_k \Bigl(\Ocircuit_{k_l}(\xvars_{k_l}) \otimes \Ocircuit_{k_r}(\xvars_{k_r}) \Bigr)
			 & \text{else},
		\end{cases}
		\label{eq:poc:def}
	\end{align}
	where the  $E_{\xvar_k}$'s are quantum state matrices, and where the $\qop_k$ are quantum operations.
\end{definition}
Note that using the Kronecker product between $\Ocircuit_{k_l}(\xvars_{k_l})$ and  $\Ocircuit_{k_r}(\xvars_{k_r})$ is a sensible choice as it describes the joint state of both subsystems.

@@

To show that PSD circuits are a special case of \puncs we now impose the following restriction on the quantum operations $\qop_k$:
\begin{align}
	\qop_k (\Ocircuit_{k_l} \otimes \Ocircuit_{k_r} )
	 & =
	\kraus_{k} \left( \Ocircuit_{k_l}  \otimes  \Ocircuit_{k_r} \right) \kraus_{k}^*
	\label{eq:pureinternal}
\end{align}
That is, we limit the quantum operation to having  only a single pair of Kraus operators. For the quantum operation to be unital we need to have $\kraus_{k} \kraus_{k}^*=\mathbb{1}$. That is, $\kraus_{k}$ has to be semi-unitary,

@@

Effectively, this would expand the research initiated by \citet{loconte2024subtractive} for mapping out the relationships between the different structured decomposable circuits. Ideally, one would like to create an analogue to \citeauthor{darwiche2002knowledge}' knowledge compilation map \citep{darwiche2002knowledge} but for positive operator circuits and relate it to the framework of algebraic model counting \citep{kimmig2017algebraic}.
This would also allow for establishing possible separations between sum-of-squares PCs~\citep{loconte2024subtractive}, product of monotonic by SOCS~\citep[Definition 5]{loconte2024subtractive}, and inception networks~\citep{wangrelationship}.
A first discussion on these issues can also be found in \citep[Appendix A]{wangrelationship}.
